Smart Drill-Down

### 2015-10-19 TableSummarization/Paper/sampling.tex
small changes throughout -- paper now under length

Leaves of tree $U$ correspond to rules that may be drilled down on next. We assume there is a probability distribution over leaves, which assigns a probability that each leaf may be drilled down on next. This can be a uniform distribution, or a machine learned distribution using past user data. We aim to set sample sizes $n_r$ so as to maximize the probability that the next drill down can be performed without accessing the hard disk.

If $r'$ is a sub-rule of $r$, and covers $x$ times as many rules as $r$, then it means that when drilling down on $r$, $s_{r'}$ can contribute around $\frac{n_{r'}}{x}$ tuples to the sample for $r$. We denote the ratio of selectivities $x$ by $S(r',r)$. $S(r',r)$ is defined to be $0$ if $r'$ is not a sub-rule of $r$. If $r$ is to be drilled down on next, the total number of sample tuples we will have for $r$ from all existing samples is given by $ess(r) = \sum_{r'\in U} S(r', r)n_{r'}$. We can drill down on $r$ without accessing hard disk, if $ess(r) \geq minSS$. We now formally define our problem: 
\begin{problem}\label{prob:sample-sizes}
Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer $M$, and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximize}$ :
$$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{[ess(r^{\prime}) \geq minSS]}$$
where the $I$'s are indicator variables, subject to :
$\sum_{r \in U} n_r \leq M$
\end{problem}
Problem~\ref{prob:sample-sizes} is non-linear and non-convex because of the indicator variables. 
%If the tree $U$ is not too big, then it might be feasible to use an exponential algorithm to solve the problem. 
We can show that Problem~\ref{prob:sample-sizes} is {\sc NP-Hard} using a reduction from the knapsack problem. 
\begin{lemma}\label{lemma:sampling-hardness}
Problem~\ref{prob:sample-sizes} is {\sc NP-Hard}.
\end{lemma}
\begin{proof}(Sketch)
Suppose we are given an instance of the knapsack problem with $m$ objects, with the $i^{th}$ object having weight $w_i$ and value $v_i$. We are also given a weight limit $W$, and our objective is to choose a set of objects that maximizes value and has total weight $< W$. We will reduce this instance to an instance of Problem~\ref{prob:sample-sizes}.

We first scale the $w_i$s and $W$ such that all $w_i$s are $< 1$. For Problem~\ref{prob:sample-sizes}, we set $M$ to $(m + W) \times minSS$. Tree $U$ has $m$ special nodes $r_1, r_2, ... r_m$, and each $r_i$ has two leaf children $r_{i,1}, r_{i,2}$. All other leaves have expansion probability $0$. 
The $S$ values are such that $\forall i \in \left\lbrace 1, 2, ..m\right\rbrace, j \in \left\lbrace 1, 2 \right\rbrace : (S(x, r_{i,j}) \neq 0 \Rightarrow x = r_{i,j} || x = r_i)$. In reality, the $S$ values cannot be exactly zero, but can be made small enough for all practical purposes. Thus, $\forall 1 \leq i \leq m, j \in \left\lbrace 1,2 \right\rbrace : ess(r_{i,j}) = n_{r_{i,j}} + n_{r_i}S(r_i, r_{i,j})$. In addition, $S(r_i, r_{i,1}) = 1$, and $S(r_i, r_{i,2}) = 1 - w_i$. Finally, $\forall i : p_{r_{i,1}} = \frac{2}{2m+1}$ and $p_{r_{i,2}} = \frac{v_i}{(2m+1)\sum_{j=1}^{m}v_i}$. So in any optimal solution, $\forall i : ess(r_{i,1}) = minSS$, and we've to decide which $i$'s also have $ess(r_{i,2}) = minSS$. Having $ess(r_{i,2}) = minSS$ requires consuming $w_i \times minSS$ extra memory and gives an extra $\frac{v_i}{(2m+1)(\sum_{j=1}^{m}v_j)}$ probability value. Thus, having $ess(r_{i,2}) = minSS$ is equivalent to picking object $i$ from the knapsack problem. Solving Problem~\ref{prob:sample-sizes} with the above $U$, $S$, $p$ and picking the set of $i$'s for which $ess(r_{i,2}) = minSS$ gives a solution to the instance of the knapsack problem. 
\end{proof}
\vspace{-10pt}
\noindent A more detailed proof sketch is given in the technical report.}

@@

We wish to set the sample size $n_r$ of each rule so as to maximize the probability that we can respond to the next user expansion without making another pass. We now formally define our problem below:
\begin{problem}\label{prob:sample-sizes}
Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer $M$, and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximize}$ :
$$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{[ess(r^{\prime}) \geq minSS]}$$
where the $I$'s are indicator variables, subject to :
$\sum_{r \in U} n_r \leq M$
\end{problem}
Problem~\ref{prob:sample-sizes} is non-linear and non-convex because of the indicator variables. 
%If the tree $U$ is not too big, then it might be feasible to use an exponential algorithm to solve the problem. 
We can show that Problem~\ref{prob:sample-sizes} is {\sc NP-Hard} using a reduction from the knapsack problem. 
\begin{lemma}\label{lemma:sampling-hardness}
Problem~\ref{prob:sample-sizes} is {\sc NP-Hard}.
\end{lemma}
\begin{proof}(Sketch)
Suppose we are given an instance of the knapsack problem with $m$ objects, with the $i^{th}$ object having weight $w_i$ and value $v_i$. We are also given a weight limit $W$, and our objective is to choose a set of objects that maximizes value and has total weight $< W$. We will reduce this instance to an instance of Problem~\ref{prob:sample-sizes}.

@@

\subsubsection{Design of the SampleHandler}\label{sec:sample-handler}
\papertext{
We now briefly describe the design of the {\em SampleHandler}, which given
a certain memory capacity $M$, and a minimum sample size $minSS$,
creates, maintains, retrieves, and removes samples, all in
response to user interactions on the table. A detailed description, with some potential optimizations, can be found in the technical report.

At all points, the SampleHandler maintains a set of samples in memory. Each sample $s$ is represented as a triple: (a) A `filter' rule $f_s$, (b) a scaling factor $N_s$ and (c) a set $T_s$ of tuples from the table. The set $T_s$ consists of a $\frac{1}{N_s}$ uniformly sampled fraction of tuples covered by $f_s$. The scaling factor $N_s$ is used to translate the count of a rule on the sample into an estimate of the count over the entire table. The sum of $|T_s|$ over all samples $s$ is not allowed to exceed capacity $M$ at any point.

>>>>

Leaves of tree $U$ correspond to rules that may be drilled down on next. We assume there is a probability distribution over leaves, which assigns a probability that each leaf may be drilled down on next. This can be a uniform distribution, or a machine learned distribution using past user data. We aim to set sample sizes $n_r$ so as to maximize the probability that the next drill down can be performed without accessing disk.

If $r'$ is a sub-rule of $r$, and covers $x$ times as many rules as $r$, then it means that when drilling down on $r$, $s_{r'}$ can contribute around $\frac{n_{r'}}{x}$ tuples to the sample for $r$. We denote the ratio of selectivities $x$ by $S(r',r)$. $S(r',r)$ is defined to be $0$ if $r'$ is not a sub-rule of $r$. If $r$ is to be drilled down on next, the total number of sample tuples we will have for $r$ from all existing samples is given by $ess(r) = \sum_{r'\in U} S(r', r)n_{r'}$. We can drill down on $r$ without accessing hard disk, if $ess(r) \geq minSS$. We now formally define our problem: 
\begin{problem}\label{prob:sample-sizes}
Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer $M$, and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximize}$ :
$$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{[ess(r^{\prime}) \geq minSS]}$$
where the $I$'s are indicator variables, with :
$\sum_{r \in U} n_r \leq M$
\end{problem}
Problem~\ref{prob:sample-sizes} is non-linear and non-convex because of the indicator variables. 
%If the tree $U$ is not too big, then it might be feasible to use an exponential algorithm to solve the problem. 
We can show that Problem~\ref{prob:sample-sizes} is {\sc NP-Hard} using a reduction from the knapsack problem. 
\begin{lemma}\label{lemma:sampling-hardness}
Problem~\ref{prob:sample-sizes} is {\sc NP-Hard}.
\end{lemma}
\begin{proof}(Sketch; details in~\cite{DBLP:journals/pvldb/JoglekarGP15})
Suppose we are given an instance of the knapsack problem with $m$ objects, with the $i^{th}$ object having weight $w_i$ and value $v_i$. We are also given a weight limit $W$, and our objective is to choose a set of objects that maximizes value and has total weight $< W$. We will reduce this instance to an instance of Problem~\ref{prob:sample-sizes}.

We first scale the $w_i$s and $W$ such that all $w_i$s are $< 1$. For Problem~\ref{prob:sample-sizes}, we set $M$ to $(m + W) \times minSS$. Tree $U$ has $m$ special nodes $r_1, r_2, ... r_m$, and each $r_i$ has two leaf children $r_{i,1}, r_{i,2}$. All other leaves have expansion probability $0$. 
The $S$ values are such that $\forall i \in \left\lbrace 1, 2, ..m\right\rbrace, j \in \left\lbrace 1, 2 \right\rbrace : (S(x, r_{i,j}) \neq 0 \Rightarrow x = r_{i,j} || x = r_i)$. In reality, the $S$ values cannot be exactly zero, but can be made small enough for all practical purposes. Thus, $\forall 1 \leq i \leq m, j \in \left\lbrace 1,2 \right\rbrace : ess(r_{i,j}) = n_{r_{i,j}} + n_{r_i}S(r_i, r_{i,j})$. In addition, $S(r_i, r_{i,1}) = 1$, and $S(r_i, r_{i,2}) = 1 - w_i$. Finally, $\forall i : p_{r_{i,1}} = \frac{2}{2m+1}$ and $p_{r_{i,2}} = \frac{v_i}{(2m+1)\sum_{j=1}^{m}v_i}$. So in any optimal solution, $\forall i : ess(r_{i,1}) = minSS$, and we've to decide which $i$'s also have $ess(r_{i,2}) = minSS$. Having $ess(r_{i,2}) = minSS$ requires consuming $w_i \times minSS$ extra memory and gives an extra $\frac{v_i}{(2m+1)(\sum_{j=1}^{m}v_j)}$ probability value. Thus, having $ess(r_{i,2}) = minSS$ is equivalent to picking object $i$ from the knapsack problem. Solving Problem~\ref{prob:sample-sizes} with the above $U$, $S$, $p$ and picking the set of $i$'s for which $ess(r_{i,2}) = minSS$ gives a solution to the instance of the knapsack problem. 
\end{proof}
%\vspace{-5pt}
%\noindent A more detailed proof sketch is given in the technical report.
}

@@

We wish to set the sample size $n_r$ of each rule so as to maximize the probability that we can respond to the next user expansion without making another pass. We now formally define our problem below:
\begin{problem}\label{prob:sample-sizes}
Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer $M$, and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximize}$ :
$$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{[ess(r^{\prime}) \geq minSS]}$$
where the $I$'s are indicator variables, subject to
$\sum_{r \in U} n_r \leq M$
\end{problem}
Problem~\ref{prob:sample-sizes} is non-linear and non-convex because of the indicator variables. 
%If the tree $U$ is not too big, then it might be feasible to use an exponential algorithm to solve the problem. 
We can show that Problem~\ref{prob:sample-sizes} is {\sc NP-Hard} using a reduction from the knapsack problem. 
\begin{lemma}\label{lemma:sampling-hardness}
Problem~\ref{prob:sample-sizes} is {\sc NP-Hard}.
\end{lemma}
\begin{proof}(Sketch)
Suppose we are given an instance of the knapsack problem with $m$ objects, with the $i^{th}$ object having weight $w_i$ and value $v_i$. We are also given a weight limit $W$, and our objective is to choose a set of objects that maximizes value and has total weight $< W$. We will reduce this instance to an instance of Problem~\ref{prob:sample-sizes}.

@@

\subsubsection{Design of the SampleHandler}\label{sec:sample-handler}
\papertext{
We now briefly describe the design of the {\em SampleHandler}, which given
a certain memory capacity $M$, and a minimum sample size $minSS$,
creates, maintains, retrieves, and removes samples, all in
response to user interactions on the table. A detailed description, with some potential optimizations, can be found in~\cite{DBLP:journals/pvldb/JoglekarGP15}.

At all points, the SampleHandler maintains a set of samples in memory. Each sample $s$ is represented as a triple: (a) A `filter' rule $f_s$, (b) a scaling factor $N_s$ (c) a set $T_s$ of tuples from the table. The set $T_s$ consists of a $\frac{1}{N_s}$ uniformly sampled fraction of tuples covered by $f_s$. The scaling factor $N_s$ is used to translate the count of a rule on the sample into an estimate of the count over the entire table. The sum of $|T_s|$ over all samples $s$ is not allowed to exceed capacity $M$ at any time.
